PhD Title \\ PhD Thesis

### 2024-07-16 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the PhD was to investigate 
the possibility of using optical tweezing as a means of initiating and 
controlling nucleation by generating fluid flow within a small droplet of 
supersaturated solution. The goal of which would be to understand the 
influence of shearing on nucleation at a micro level as compared to larger 
scale results. It has been shown that for macro-scale systems, the likelihood 
of nucleation increases to a maximum value under increased shearing 
\cite{Debuysschere2023, Mura2016}. Theoretical research into the matter 
identified two competing effects that effect a crystal in moving fluid fields; 
firstly, nucleation is enhanced due to the increased mass transfer of solute 
material; and secondly, shear flow against the crystal surface leading to a 
decrease in growth \cite{Mura2016}. These two competing effects are validated 
by experimental work using glycine solutions, showing that beyond a certain 
shear rate the nucleation rate is reduced \cite{Debuysschere2023}. In this 
chapter I outline the optical tweezer equipment used during this PhD, the 
initial attempts made to induce nucleation via optical rotation, and the 
impact of a moving beam on the crystal growth of a newly formed nucleus.

\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, two
microscope objectives (one for trapping and one for imaging),
and a means of controlling the position of the loaded sample. 
While there are other pieces of equipment used in modern tweezer
experiments these are the bare minimum requirements for any 
optical tweezer. The laser used for this project was a 1064 
nm near infrared laser - provided by CNI Lasers – that has 
an adjustable power supply to vary the energy output of the 
laser - the remaining optics where supplied by Thor Labs.
Experimental work has shown that the trapping efficiency 
increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ 
\cite{kim2003dependence} where $D_{obj}$ is the diameter
of the objective aperture. To expand the beam front we 
utilise a Galilean beam expansion arrangement (indicated 
by $f_1$, and $f_2$) as recommended for high power laser 
applications. In our initial experiments the beam expansion 
provides a $4\times$ magnification; however, in later experiments 
where we utilise a galvano-mirror the beam expansion is $3\times$ 
and then the 4f correlator provides a further $1.25\times$ magnification 
(using $f_3$ and $f_4$) - where the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}
It should be noted that the galvano-mirror requires the use 
of a Keplerian beam expansion arrangement which reduces
the transmitted laser power due to localised heating of the air.
Afterwards the laser is passed through a dichroic mirror that 
separates incoming infrared and visible light, this is to 
prevent the laser from damaging the CCD camera used for imaging
the trapping plane. The laser is then focused to a diffraction
limited spot by a 1.25 NA objective. By increasing the numerical
aperture of the objective, the gradient force at the focal point 
is increased; the trade-off being that the for higher NA values 
the trapping depth is reduced due to spherical aberrations.
While it is possible to increase the trapping depth \cite{Reihani2006} 
by adjusting the objective's tube length this approach is incompatible 
with our trapping arrangement. A 0.25 NA condenser objective refocuses
the scattered laser light and also provide an aperture for an 
imaging LED to illuminate the focal plane. Samples are loaded 
onto a piezo driven table to that is inserted between the trapping
and condensing objectives; the piezo drivers allow for sub-micron
control of the sample position to a degree as small as a 10 nm. 
To detect and monitor the position of a trapped particle a 
quadrant photo diode (QPD) was utilised.

\begin{figure}
	\label{fig:setup}
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the PhD. The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively. Diagram not drawn to scale.}
\end{figure}

\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, 
a position detection system must be utilised. There are 3 possible
methods of position detection: video-analysis, lateral-effect
position sensing, and quadrant photodiodes. The former being 
ideally suited for multiple traps or situations where precision
is not the top priority. In order to match the force measurements
of back-focal plane interferometry requires the camera's frame
rate to exceed $1\ kHz$ which can be difficult to achieve while 
maintaining a decent resolution \cite{Gibson2008}. In comparison
off the shelf back-focal plane detectors can achieve temporal
resolutions anywhere from $10-100\ kHz$.

A QPD is frequently used position detection system for optical 
tweezers due to their high sampling rate, high degree of 
precision, and ease of set up. The QPD is constructed of four photo 
diodes assembled in a quadrant formation, when a particle is trapped the 
interference pattern produced is focused onto the QPD, with 
the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants 
together the particle's centre of mass is tracked in the x-y 
plane. Axial displacement can be estimated by observing the change
in the total voltage of the QPD. The outputted signal gives an
indication of the particle's relative displacement from the beam 
focus, but in order to convert the signal to distance units the
trap needs to be calibrated (assuming a linear response curve).

\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
\end{figure}

@@

\subsection{Fourier Optics and 4f correlators}
A 4f correlator is an example of Fourier optics in practice, understanding
that a focused lens takes a Fourier transform of the light profile. 
Consider a laser with a circular Gaussian profile, if you were to place 
a detector there you would pick up the intensity as a function of its 
position within the beam. If however you focused the light into a single 
point (using a +ve focal lens) you are actually seeing a measurement of 
the phase of your laser with position, in which you would see a diffraction
limited spot ($d = \lambda/2nsin(\theta)$), indicating that the laser is 
collimated. In imaging systems, a series of focal lenses can be used to 
filter out unwanted scattering from an image (or in an inverse case 
differentiate between different images), the placement of each lens is shown below.
For our applications a 4f correlator is utilised to ensure that the motion
of the galvano-mirrors does not move the focal point of the laser, allowing 
for a stable trap even while in motion.

\section{Calibration of Tweezer Setup}
Prior to testing optical rotation the trapping laser was calibrated 
using power spectra analysis

\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
Generation of fluid shear can be achieved via two avenues: Firstly, by 
utilising circularly polarised light it is possible to transfer angular 
momentum from the laser to the trapped entity. Secondly, one can 
directly move the trap within the imaging plane by steering the beam 
using either a galvanometric mirror or gimbal mirror. The following 
chapter outlines the work done with shear generated by circularly 
polarised light and the challenges of applying this to localising 
nucleation.

@@

Vaterite samples where made by the first preparing equal amounts of 
$CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at the same time
a vial of $0.33M\ MgSO_4$ was prepared and set aside for later. First
a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of Agepon was added
to halt the reaction, the solution was filtered and washed 3 times with 
distilled water before being suspended in water. When trapped in circularly 
polarised light, the anisotropic scattering of the sphere results in a periodic 
signal on the QPD. Therefore, the resulting power spectrum is not a Lorentzian
but now also displays peaks that appear at integer multiples of the particles
rotational frequency.

\begin{figure}[h!]
	\label{fig:vaterite}
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap. (b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
\end{figure}

@@

\subsection{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
Using a birefringent particle and rotating it within the fluid, the maximum 
rotation rate is due to the fluid drag resisting the torque of the 
trapping beam \cite{RodriguezSevilla2018}. If you want to measure 
fluid flow you can instead use a micro-rotor to see how fluid flow 
propagates in the medium \cite{Knoener2005}. Likewise, one can use 
a galvanometric mirror to probe the drag force of the fluid, by 
understanding the trap strength (calibrating using a low frequency 
signal) one can measure the drag force experienced by the local fluid \cite{RobertsonAnderson2018}. I

Understanding the fluid velocity around our trapped object is determined 
mostly by the Reynold's number of our system, for a sphere submersed 
in a moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}
Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's 
density and viscosity respectively. In our case we do not have a fluid
moving around a sphere but a sphere moving through the fluid at some 
velocity $U$, assuming a no-slip boundary condition we can model the 
fluid velocity profile based on the velocity of the particle. There are 
two possible avenues for generating shear flow with a trapped particle; 
rotation of birefringent particles, and fluid flow induced by particle motion.

@@

First we tested the rotational behaviour of vaterite in distilled water, 
samples of vaterite where diluted down and $200\mu L$ was pipetted onto
the sample stage. Due to Van der Waal's forces some of the microspheres 
were stuck together, fortunately individual sphere's were still present.
Over the course of a few weeks multiple sphere's where trapped 
and their rotation speeds were measured via the QPD.

\begin{figure}[h!]
	\label{fig:vaterite_shear}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere. (b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line. Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW. Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
\end{figure}

@@

\section{Rotation of micro-rotors in supersaturated solutions}
Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of 
supersaturated solutions of Glycine and water. After locating a 
singular microsphere the vaterite was trapped in a circularly 
polarised light and brought close to the droplet edge, after a 
period of ten minutes if no nucleation event was observed the 
particle was released and another particle was trapped. Due to 
the increased viscosity of the supersaturated solution and the 
proximity to the droplet edge, this was done as previous work with
optical tweezers have only seen nucleation near the solution-air interface
\cite{Gowayed2021, Liao2022, Walton2018}. The observed rotation rate 
was rather low or non-existent, the results are tabulated below:
\begin{table}[h!]
	\centering
	\caption{Results from rotating vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.29} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.40} & 6.32 & 0.00 & $\times$ \\
		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}

@@

\section{Shearing via a Galvano-mirror}
Rather than utilising birefringence rotation, which is subject to issues
during synthesis and achieving an ideal particle size, a galvano-mirror
was installed to induce controlled particle motion. While typically 
galvano and gimble mirrors are used to trap multiple particles in a 
regular pattern, in a sufficiently dilute solution a single silica 
micro sphere can be moved quickly through a fluid along a preset path.
Calculating the shear rate around an individual particle is difficult to
do precisely but for low Reynolds numbers we can get an adequate approximation.

@@

\section{Nucleation using a moving beam}
As mentioned previously, shearing via optical rotation and particle displacement
did not result in

>>>>

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the PhD was to investigate 
the possibility of using optical tweezing as a means of initiating and 
controlling nucleation by generating fluid flow within a small droplet of 
supersaturated solution. The goal of which would be to understand the 
influence of shearing on nucleation at a micro level as compared to larger 
scale results. It has been shown that for macro-scale systems, the likelihood 
of nucleation increases to a maximum value under increased shearing 
\cite{Debuysschere2023, Mura2016}. Theoretical research into the matter 
identified two competing effects that effect a crystal in moving fluid fields; 
firstly, nucleation is enhanced due to the increased mass transfer of solute 
material; and secondly, shear flow against the crystal surface leading to a 
decrease in growth \cite{Mura2016}. These two competing effects are validated 
by experimental work using glycine solutions, showing that beyond a certain 
shear rate the nucleation rate is reduced \cite{Debuysschere2023}. In this 
chapter I outline the optical tweezer equipment used during this PhD, the 
initial attempts made to induce nucleation via optical rotation, and the 
impact of a moving beam on the crystal growth of a newly formed nucleus.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, two
microscope objectives (one for trapping and one for imaging),
and a means of controlling the position of the loaded sample. 
While there are other pieces of equipment used in modern tweezer
experiments these are the bare minimum requirements for any 
optical tweezer. The laser used for this project was a 1064 
nm near infrared laser - provided by CNI Lasers – that has 
an adjustable power supply to vary the energy output of the 
laser - the remaining optics where supplied by Thor Labs.
Experimental work has shown that the trapping efficiency 
increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ 
\cite{kim2003dependence} where $D_{obj}$ is the diameter
of the objective aperture. To expand the beam front we 
utilise a Galilean beam expansion arrangement (indicated 
by $f_1$, and $f_2$) as recommended for high power laser 
applications. In our initial experiments the beam expansion 
provides a $4\times$ magnification; however, in later experiments 
where we utilise a galvano-mirror the beam expansion is $3\times$ 
and then the 4f correlator provides a further $1.25\times$ magnification 
(using $f_3$ and $f_4$) - where the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

It should be noted that the galvano-mirror requires the use 
of a Keplerian beam expansion arrangement which reduces
the transmitted laser power due to localised heating of the air.
Afterwards the laser is passed through a dichroic mirror that 
separates incoming infrared and visible light, this is to 
prevent the laser from damaging the CCD camera used for imaging
the trapping plane. The laser is then focused to a diffraction
limited spot by a 1.25 NA objective. By increasing the numerical
aperture of the objective, the gradient force at the focal point 
is increased; the trade-off being that the for higher NA values 
the trapping depth is reduced due to spherical aberrations.
While it is possible to increase the trapping depth \cite{Reihani2006} 
by adjusting the objective's tube length this approach is incompatible 
with our trapping arrangement. A 0.25 NA condenser objective refocuses
the scattered laser light and also provide an aperture for an 
imaging LED to illuminate the focal plane. Samples are loaded 
onto a piezo driven table to that is inserted between the trapping
and condensing objectives; the piezo drivers allow for sub-micron
control of the sample position to a degree as small as a 10 nm. 
To detect and monitor the position of a trapped particle a 
quadrant photo diode (QPD) was utilised.

\begin{figure}
	\label{fig:setup}
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the PhD. The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively. Diagram not drawn to scale.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, 
a position detection system must be utilised. There are 3 possible
methods of position detection: video-analysis, lateral-effect
position sensing, and quadrant photodiodes. The former being 
ideally suited for multiple traps or situations where precision
is not the top priority. In order to match the force measurements
of back-focal plane interferometry requires the camera's frame
rate to exceed $1\ kHz$ which can be difficult to achieve while 
maintaining a decent resolution \cite{Gibson2008}. In comparison
off the shelf back-focal plane detectors can achieve temporal
resolutions anywhere from $10-100\ kHz$.

A QPD is frequently used position detection system for optical 
tweezers due to their high sampling rate, high degree of 
precision, and ease of set up. The QPD is constructed of four photo 
diodes assembled in a quadrant formation, when a particle is trapped the 
interference pattern produced is focused onto the QPD, with 
the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants 
together the particle's centre of mass is tracked in the x-y 
plane. Axial displacement can be estimated by observing the change
in the total voltage of the QPD. The outputted signal gives an
indication of the particle's relative displacement from the beam 
focus, but in order to convert the signal to distance units the
trap needs to be calibrated (assuming a linear response curve).
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
A 4f correlator is an example of Fourier optics in practice, understanding
that a focused lens takes a Fourier transform of the light profile. 
Consider a laser with a circular Gaussian profile, if you were to place 
a detector there you would pick up the intensity as a function of its 
position within the beam. If however you focused the light into a single 
point (using a +ve focal lens) you are actually seeing a measurement of 
the phase of your laser with position, in which you would see a diffraction
limited spot ($d = \lambda/2nsin(\theta)$), indicating that the laser is 
collimated. In imaging systems, a series of focal lenses can be used to 
filter out unwanted scattering from an image (or in an inverse case 
differentiate between different images), the placement of each lens is shown below.
For our applications a 4f correlator is utilised to ensure that the motion
of the galvano-mirrors does not move the focal point of the laser, allowing 
for a stable trap even while in motion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Calibration of Tweezer Setup}
Prior to testing optical rotation the trapping laser was calibrated 
using power spectra analysis

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
Generation of fluid shear can be achieved via two avenues: Firstly, by 
utilising circularly polarised light it is possible to transfer angular 
momentum from the laser to the trapped entity. Secondly, one can 
directly move the trap within the imaging plane by steering the beam 
using either a galvanometric mirror or gimbal mirror. The following 
chapter outlines the work done with shear generated by circularly 
polarised light and the challenges of applying this to localising 
nucleation.

@@

Vaterite samples where made by the first preparing equal amounts of 
$CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at the same time
a vial of $0.33M\ MgSO_4$ was prepared and set aside for later. First
a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of Agepon was added
to halt the reaction, the solution was filtered and washed 3 times with 
distilled water before being suspended in water. When trapped in circularly 
polarised light, the anisotropic scattering of the sphere results in a periodic 
signal on the QPD. Therefore, the resulting power spectrum is not a Lorentzian
but now also displays peaks that appear at integer multiples of the particles
rotational frequency. 
\begin{figure}[h!]
	\label{fig:vaterite}
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap. (b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
Using a birefringent particle and rotating it within the fluid, the maximum 
rotation rate is due to the fluid drag resisting the torque of the 
trapping beam \cite{RodriguezSevilla2018}. If you want to measure 
fluid flow you can instead use a micro-rotor to see how fluid flow 
propagates in the medium \cite{Knoener2005}. Likewise, one can use 
a galvanometric mirror to probe the drag force of the fluid, by 
understanding the trap strength (calibrating using a low frequency 
signal) one can measure the drag force experienced by the local fluid \cite{RobertsonAnderson2018}. I

Understanding the fluid velocity around our trapped object is determined 
mostly by the Reynold's number of our system, for a sphere submersed 
in a moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}

Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's 
density and viscosity respectively. In our case we do not have a fluid
moving around a sphere but a sphere moving through the fluid at some 
velocity $U$, assuming a no-slip boundary condition we can model the 
fluid velocity profile based on the velocity of the particle. There are 
two possible avenues for generating shear flow with a trapped particle; 
rotation of birefringent particles, and fluid flow induced by particle motion.

@@

First we tested the rotational behaviour of vaterite in distilled water, 
samples of vaterite where diluted down and $200\mu L$ was pipetted onto
the sample stage. Due to Van der Waal's forces some of the microspheres 
were stuck together, fortunately individual sphere's were still present.
Over the course of a few weeks multiple sphere's where trapped 
and their rotation speeds were measured via the QPD. 
\begin{figure}[h!]
	\label{fig:vaterite_shear}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere. (b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line. Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW. Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation of micro-rotors in supersaturated solutions}
Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
(see sec.~\ref{sec:vaterite}), and then suspended in distilled 
water, at the same time a supersaturated solution of Glycine 
and water was prepared and $50\ \mu L$ was pipetted onto a 
glass cover slip. A single microsphere the vaterite was trapped 
in a circularly polarised light and brought close to the droplet 
edge.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
	\caption{Diagram of optical trapping set up for rotating birefringent 
		particles in a supersaturated solution. Left: side view of the 
		trapping set up showing the location of the trap focus at the 
		edge of the droplet of a supersaturated solution. Right: top down 
		view of the glycine droplet with a trapped birefringent particle 
		shown close to edge of the trap. As the particle rotates the drag 
		force from the surrounding fluid generates a flow field around 
		itself (see Eq.~\ref{eq:birefringent_speed}).}
\end{figure}

After a period of ten minutes if no nucleation event was 
observed the particle was released and another particle was 
trapped. Due to the increased viscosity of the supersaturated 
solution and the proximity to the droplet edge, as previous works
have only seen nucleation near the solution-air interface \cite{Gowayed2021, 
Liao2022, Walton2018}, the observed rotation rate was rather low or 
non-existent, the results are tabulated below:
\begin{table}[h!]
	\centering
	\caption{Results from rotating vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.4} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.45} & 6.32 & 0.00 & $\times$ \\
		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Shearing via a Galvano-mirror}
Rather than utilising birefringence rotation, which is subject to issues
during synthesis and achieving an ideal particle size, a galvano-mirror
was installed to induce controlled particle motion. While typically 
galvano and gimble mirrors are used to trap multiple particles in a 
regular pattern, in a sufficiently dilute solution a single silica 
micro sphere can be moved quickly through a fluid along a preset path.
Calculating the shear rate around an individual particle is difficult to
do precisely but for low Reynolds numbers we can get an adequate approximation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nucleation using a moving beam}
As mentioned previously, shearing via optical rotation and particle displacement
did not result in any localised nucleation events even while in the proximity of 
the droplet edge. During the experiments with the galvano-mirror, it was found 
that when no particle was present in the optical trap nucleation events would 
occur while the beam was close to the edge of the droplet, even though the solution
was undersaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
but was more interesting is how the beam's motion influenced the growth of the nucleus.

Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
event in supersaturated glycine solution (S=1.03), the beam is a stationary being 
$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly until after roughly $6$ seconds 
the crystal escapes the trap. The key take away to remember is that the beam has no 
real influence over the crystal front,instead the crystal grows outward from the 
centre of the focus.
\begin{figure}[h!]
	\label{fig:stationary_beam}
	\centering
	\includegraphics[width=0.7\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape the
	trap.}
\end{figure}

Now while trying to induce fluid flow by optically trapping a silica particle and 
using a galvano-mirror to move the particle through the solution, nucleation events
were observed in which the crystal front follows the path of the laser. See below 
in \textcolor{red}{Fig.XX} where we have the laser focus moving in a small elliptical 
pattern,

### 2024-07-16 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
\label{chapter:simulated_detection}
\section{Simulative QPD}
A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.

>>>>

\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
\label{chapter:simulated_detection}
As outlined in the end of Chapter 4, one of the difficulties in characterising 
interactions with asymmetric objects is the coupled motion between translation 
and rotation. This can be somewhat ignored if the rotational motion is periodic,
like when the power spectra of elliptical polystyrene particles was fitted by 
Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, however in the case where 
rotational motion is stochastic the problem is more complex. For example, when a
novel trap characterisation technique was implemented by Saffron \textit{et al}
\cite{BarZiv1997, Meller1998}, they were able to use dynamic light scattering to
characterise both the axial and lateral trap stiffness acting on microspheres. The
only drawback admitted to in their work was that the technique was constrained to
isotropic scatters as their theoretical model for describing the auto-correlation
function was predicated on the fact that any variations in the signal are due to
the particles translational motion within the confines of a cylindrical trap.

\section{Simulative QPD}
A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
